---
title: "stats.project"
author: "Adam Grottoli"
date: "March 6, 2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/Adam/Documents/grad_school/Broders/Data/stats.project")
library(tidyr)
library(reshape2)
library(ggplot2)
library(bestNormalize)
```

```{r, results=FALSE}
all.vary=read.csv('all.vary.csv') ## Read data file with intraspecific variation calculated and duplicate measurement from bats removed
pairs <- t(combn(nrow(all.vary), 2)) 
pairs <- pairs[all.vary$bat[pairs[,1]] != all.vary$bat[pairs[,2]],]##create all pssible pairs of individuals
```

```{r, results=FALSE}
selected.pairs <- sample(nrow(pairs), 213, replace=TRUE)
(samps <- pairs[selected.pairs,])## Randomly select 213 pairs of individuals
```

```{r, results=FALSE}
(obs1 <- all.vary[pairs[selected.pairs,1],])
```

```{r, results=FALSE}
(obs2 <- all.vary[pairs[selected.pairs,2],]) ##Create 2 new dfs for each pair of observations
```


```{r}
obs2$fa2<-obs2$fa
fa.dif<-data.frame(obs1$bat, obs1$fa, obs2$bat, obs2$fa2) ## Merge measurements from pairs and assocciated bat ID one df 
fa.dif$inter.dif<-abs(fa.dif$obs1.fa-fa.dif$obs2.fa2) ##Find absolute difference between measurements from each pair
```

```{r}
fa.dif$intra.dif<-all.vary$intra.vary ## Merge intra-individual with inter-individual data
```

```{r}
melt.fa=melt(fa.dif, measure.vars=c('inter.dif', 'intra.dif')) ## Melt data so it can be read by ggplot
  ggplot(melt.fa, aes(x = variable, y = value)) +
  geom_boxplot() +
  scale_x_discrete (labels=c("Interindividual", "Intraindividual")) +
  labs(y='Difference in Forearm Length Measurements (mm)', x='Variation Type') +
  theme_classic()## Visualise data with ggplot
```
Figure 1. Comparing the interindividual FA measurement variation with the intraindividual FA measurement variation.

```{r}
ggplot(data=melt.fa, aes(x=value)) +
geom_histogram(binwidth = 0.1) +
labs(x='FA Variation (mm)', y='Frequency') +
theme_classic() ## Check distribution of data
```
Figure 2. Frequency of the of the FA variation including both inter- and intra-individual variation.

```{r}
shapiro.test(melt.fa$value) ## Test for normality
```

```{r}
melt.fa$log.fa=log(melt.fa$value)
```

```{r}
bestNormalize(melt.fa$value)
melt.fa$cbrt.fa=sign(melt.fa$value) * abs(melt.fa$value)^(1/3) ## Data is not normal (right skew) so it does not meet assumtions of the ANOVA. Transform data using a cubed root transformation.
ggplot(data=melt.fa, aes(x=cbrt.fa)) +
geom_histogram(binwidth = 0.1) +
labs(x='Cubed Root FA Variation', y='Frequency') +
theme_classic() ## Data looks better following transformation

```
Figure 3. Frequency of FA measurement variation that has undergone a cubed root transformation.
```{r}
shapiro.test(melt.fa$cbrt.fa) ## Test for normality again
```


```{r}
wilcox.test(melt.fa$value)## Still not normal so we will use a non-parametric test
```


```{r}
fa.dif$intra.cube.rt=sign(fa.dif$intra.dif) * abs(fa.dif$intra.dif)^(1/3) ## Transform intra-individual data to make normal
shapiro.test(fa.dif$intra.cube.rt) ## Fails the Shapiro test
g1 = fa.dif$intra.cube.rt
m1<-mean(g1)
std1<-sqrt(var(g1))
hist(g1, density=20, breaks=20, prob=TRUE, 
     xlab="Cubed Root Measurement Variation", ylim=c(0, 6), main=NULL)
curve(dnorm(x, mean=m1, sd=std1), 
      col="darkblue", lwd=2, add=TRUE, yaxt="n") ## Data still follows a normal distribution relatively well
```
Figure 4. Intra-individual FA measurement variation that has undergone a cubed root tranformation with overlayed a normal disribution curve (mean=o.693, sd=0.211)

```{r}
.5+(1-(-2*(pt(.5^(1/3), 212)-1)))/2 ## Using a t distribution we can estimate the average probability any given measurement will have an error of +/- 0.5 mm. We use the cubed root of 0.5 to match the transformation.
```

```{r}
fa_mass=read.csv('fa_mass.csv')
bat_mass=read.csv('fa_mass.csv')
ggplot()+ 
   geom_smooth(data=bat_mass, method='glm',aes(x=bat_mass$fa, y=bat_mass$mass))+
   geom_point(data=bat_mass, aes(x=bat_mass$fa, y=bat_mass$mass))+
labs(x='Forearm Length (mm)', y='Mass (g)')
fit.mass_fa=glm(bat_mass$mass~bat_mass$fa)
summary.lm(fit.mass_fa)
```

```{r}
x1 <- 32.49:41.74

b0 <- -9.34288
b1 <- 0.44976

sigma <- 0.211

eps <- (rnorm(x1,.693,sigma))^3
y <- b0 + b1*x1 + eps
summary(lm(y~x1))
plot(y)
```

fa.dif$inter.cube.rt=sign(fa.dif$inter.dif) * abs(fa.dif$inter.dif)^(1/3)